### 2015-08-30 paper/one/summary.tex
Added summary for first paper
I used LyX to generate the TeX code. Let me know if you're having
problems reading/viewing it.

%% LyX 2.1.4 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{babel}
\begin{document}

\title{Paper One Summary}

\author{Matthew Neal, Joseph Sankar, and Sasha Sobran}

\maketitle

\section*{Reference}

Daryl Posnett, Vladimir Filkov, and Premkumar Devanbu. 2011. Ecological
Inference in Empirical Software Engineering. Proceedings of the 2011
26th IEEE/ACM International Conference on Automated Software Engineering.

\section*{Important Keywords}

\section*{Feature Extraction}

\section*{Possible Improvements}
\end{document}