### 2015-09-03 hw/read/1/summary.tex
Added two possible improvements to paper one

%% LyX 2.1.4 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{babel}
\begin{document}

\title{Paper One Summary}

\author{Matthew Neal, Joseph Sankar, and Alexander Sobran}

\maketitle

\section*{Reference}

Daryl Posnett, Vladimir Filkov, and Premkumar Devanbu. 2011. Ecological
Inference in Empirical Software Engineering. Proceedings of the 2011
26th IEEE/ACM International Conference on Automated Software Engineering.

\section*{Important Keywords}
\begin{description}
\item [{Ecological~inference}] The idea that an observation or finding
at an aggregated level can apply at a disaggregated level. For example,
the idea that a model of defects at the package level also applies.
at the file level 
\item [{Ecological~fallacy}] An incorrect application of ecological inference.
A discrepency between the findings at an aggregated level and disaggragated
level. For example, an instance where a model of defects at the product
level is incorrectly used to model defects at the file level. 
\item [{Scale}] Size of an aggregated unit. The larger the scale the bigger
and fewer the aggregated units. This can effect affect the quality
of statistical models built for defect prediction. 
\item [{Zonation}] The manner in which aggregation is performed. For example,
gerrymandering. This can create internal validity threats and lead
to ecological fallacies. 
\item [{Aggregate~Phenomena}] Certain phenomena may apply only at an aggregated
level. The example given in the paper is that of inheritance. Inheritance
only takes place at the class level and cannot be studied at a disaggregated
level. 
\item [{Cost~Effectiveness}] A concept used within the paper to assess
defect models. The baseline is that funds are given to investigate
{10\%} of a project for defects, in {10\%} of the lines of code
taken at random {10\%} of the defects should be found. Models are
compared against this baseline. 
\end{description}

\section*{Feature Extraction}
\begin{itemize}
\item \textbf{Motivational Statements} Modern software applications are
composed of many files. Thus, the files are usually organized into
units, which may themselves be organized into broader units. While
research could be focused at the most granular level, for example
files, methods, even lines of code, research at these levels is costly
and takes time, so research instead usually examines data at higher
levels. Therefore, it would be useful to draw conclusions from higher
levels that would apply to lower levels. 
\item \textbf{Related Work} To improve model performance, Koru \textit{et
al.} recommend aggregating. Their methods to acheive improved performance
include summing method-level measures into class-level measures and
combining these with metrics from class-level features. They conclude
that this approach overcomes the issues of asymmetry in defect data
sets. Schröter \textit{et al.} conclude that predicting at coarser
granularities is less difficult in comparison to finer granularities
but lead to worse results. Zimmerman \textit{et al.} find different
results where package level correlations and model correlations are
higher than file level correlations and model correlations. Ambrose
\textit{et al.} argue against the use of aggregate level metrics,
though without data, stating predictions are more cumbersome to investigate,classes
are inherently self-contained, and information derived from the package-level
cannot be applied to the class level. 
\item \textbf{Motivational Statements}: The authors poses their motivation
along with the central question of the paper \textquotedbl{}What is
the right level of study?\textquotedbl{} (At an aggregated level or
dis-aggregated level). They believe that this question has a central
significance to whether or not a model that is built at a specific
level of aggregation can hold at a lower level of aggregation as well
as whether performance of the model as regards cost prediction is
valid. 
\item \textbf{New Results}: The authors grouped variables in the models
based on loss of significance(LS) or gain of significance(GS). They
explain that both LS and GS can cause Ecological Inferences issues
as in a loss of significance situation when inferring between the
aggregate level and the disaggregate level or GS when inferring from
the disaggregate to the aggregate. They found 28 instances of GS and
25 instances of LS out of 108 variables over 68 models. The authors
draw the conclusion that models built at the aggregated level should
be undertaken with the knowledge of the underlying risk of Ecological
Inference and Ecological Fallacy. 
\item \textbf{Future Work}:The authors bring up three sources of Ecological
Inference risk: Zonation, Sample Size, and Class Imbalance. The authors
believe that further study of the three sources is in order. To give
a short background on how each applies:

\begin{itemize}
\item Zonation: Aggregation runs the risk of creating distortion by the
way that the data is aggregated. 
\item Sample Size: As aggregation takes place inherently there is move to
smaller and smaller sample sizes. As the move to those smaller sample
sizes takes place there is a loss in Statistical power. 
\item Class Imbalance: A situation where a set of samples have a particular
attribute and a large proportion of that set have a particular value
with a very small number having another significant but underrepresented
value. The example in the paper is a situation where there are a very
high number of projects without defects and a very lower number with
defects. 
\end{itemize}
\item \textbf{Sampling Procedures}: Projects were taken from a set of 18
Apache Software Foundation Projects over 87 versions. The writers
took JIRA issue reports and then mapped them to pull commits. Then,
they took those commits and connected them to packages and files.
They also pulled the number of lines of code as well as the number
of developers per project. They were unable to look at past data sets
due to their inability to correctly aggregate the data given that
many data sets do not contain issue identifiers.
\end{itemize}

\section*{Possible Improvements}
\begin{itemize}
\item There are some particularly distracting typos in the paper. For example,
on pages 367-368 the authors failed to correctly reference their table
correctly leaving \textquotedbl{}Table ??\textquotedbl{} in the middle
of their text twice. 
\item The authors could have provided a better explanation of AUCROC and
AUCCE and the meaning of their results with respect to those measures.
The authors did a good job of comparing file and package measures,
but did not explain why file AUCCE was lower than file AUCROC. Is
this just the nature of AUCCE since it is including cost effectiveness?
Why was there more variance in the AUCCE measures than AUCROC?
\item At the end of the paper, the authors provide a brief note about threats
to the external validity of the results. The authors only examined
projects from one source: the Apache Software Foundation. Analyzing
projects from other sources which employ different software engineering
principles or which organize their projects differently would bolster
their results and conclusions.\end{itemize}

\end{document}