### 2015-09-07 hw/read/2/summary.tex
Initialized second paper summary and added definitions

\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{babel}
\begin{document}

\title{Paper Two Summary}

\author{Matthew Neal, Joseph Sankar, and Alexander Sobran}

\maketitle

\section*{Reference}

Erik Arisholm, Lionel C. Briand, and Eivind B. Johannessen. 2010. A systematic and comprehensive investigation of methods to build and evaluate fault prediction models. J. Syst. Softw. 83, 1 (January 2010), 2-17. DOI=10.1016/j.jss.2009.06.055 http://dx.doi.org/10.1016/j.jss.2009.06.055

\section*{Important Keywords}
\begin{description}
\item [{Fault-proneness}] The number of defects detected in a software-component. A measurement of the likelihood of software or software component error.
\item [{Object-oriented~(OO)~code~measures}] Measurement of the structural properties of the source code.
\item [{Delta~measures}] Measurement of the amount of change (churn) in a file between two successive releases.
\item [{Process~measures}] Measures from a configuration management system. Includes developer experience, the number of developers that have edited a file, the number of faults in the previous releases, the number of line added, the number of lines removed.

\end{description}

\end{document}