### 2015-09-17 hw/read/3/summary.tex
Began summary for Paper 3

\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{babel}
\begin{document}

\title{Paper Three Summary}

\author{Matthew Neal, Joseph Sankar, and Alexander Sobran}

\maketitle

\section*{Reference}

Bird et al \cite{bird09} listed below.

\section*{Important Keywords}
\begin{description}
\item [{Global development}] Development that is globally distributed but takes place within one company. The authors contrast this to \emph{outsourcing}, in which development takes place among multiple companies and \emph{collocated development}, where development takes place in one location. 
\item [{Coordination breakdown}] A difficulty usually encountered in distributed development where employees in one location are not appropriately aware or responsive to the work of employees in another location.
\item [{Binary distribution}] Different levels of classifications for where binaries are developed. The authors give \emph{building, cafeteria, campus, locality, continent, } and \emph{world} as different ways to classify binary distribution.
\item [{Organizational integration}] An organizational structure that spans geographical locations at low levels.
\end{description}

\section*{Feature Extraction}

\section*{Possible Improvements}

\section*{Connection to Other Papers}

\bibliographystyle{plain}
\bibliography{references}
\end{document}