### 2015-09-24 hw/read/4/summary.tex
initialize paper four

\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{babel}
\begin{document}

\title{Paper Four Summary}

\author{Matthew Neal, Joseph Sankar, and Alexander Sobran}

\maketitle

\section*{Reference}

Vandecruys et al \cite{Vandecruys08} listed below.

\section*{Important Keywords}
\begin{description}
\item [{Ant Colony Based Optimization}] 
\end{description}

\section*{Feature Extraction}
\begin{description}
\item [{Motivational~statements}]
\end{description}

\section*{Possible Improvements}

\section*{Connection to Other Papers}

\bibliographystyle{plain}
\bibliography{references}
\end{document}