### 2015-11-30 hw/read/9/summary.tex
Added some features for reading 9

\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{babel}
\usepackage{hyperref} 
\begin{document}
\title{Paper Nine Summary}

\author{Matthew Neal, Joseph Sankar, and Alexander Sobran}

\maketitle

\section*{Reference}

Nam et al. \cite{Nam} listed below.

\section*{Important Keywords}
\begin{description}
\item[{Transfer defect learning}] Extracting common knowledge from one task domain and transferring it to another. The transferred knowledge is then used to train a prediction model.
\item[{TCA+}] An improvement on Transfer Component Analysis. TCA can map the data of the source and target projects on a latent feature space, but is sensitive to normalization. TCA+ selects a proper normalization to yield better prediction performance.
\item[{Data set Characteristic Vector}] A vector of six elements each relating to the distance between pairs of instances of data. DCVs are used to see how similar two projects are (see below).
\item[{Similarity vector}] Represents the difference between two projects: a source and a target. Examples include "much more", "less", or "same". The values in the DCVs are used to calculate the similarity vectors.
\end{description}

\section*{Feature Extraction}
\begin{description}
\item[{Motivational statements}] Cross-project defect prediction is often necessary for new projects, but doesn't always yield good results. Therefore, the authors hope to employ transfer defect learning to improve the performance of these models.
\item[{Data}] The authors have provided the data used in their experiments at\\
 \href{https://sites.google.com/site/transferdefect/}{https://sites.google.com/site/transferdefect/}.
\item[{Future work}] The authors were looking into transferring knowledge across entire domains as an extension of transfer learning. They were also interested in seeing which other prediction and recommendation systems might benefit from transfer learning.
\end{description}

\section*{Possible Improvements}

\section*{Connection to Other Papers}

\bibliographystyle{plain}
\bibliography{references}
\end{document}